---
title: "Frito Lay: Customer Attrition"
output: html_notebook
---

```{r}
library(tidyverse)
library(caret)
library(e1071)
library(class)
library(ggthemes)
```


# ------------ Data Loading -------------
```{r}
getwd()
data <- read.csv("CaseStudy1-data.csv")
```

# ------------ Data Quality, Structure & Sanity Checks -------------
```{r}
# Data Dimensionality
dim(data) 
```

```{r}
# Missing Values
sum(is.na(data)) 
```

```{r}
str(data)
```


```{r}
# Summary statistics
summary(data)
```

```{r}
# Duplicate Analysis
sum(duplicated(data))
```

```{r}
# Columns in data
colnames(data)
```

```{r}
#Checking out the Attrition Distribution
ggplot(data = data, mapping = aes(x = Attrition))  + 
  geom_bar() + 
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.5,
    size = 4,
    color = "black"
  ) +
  theme_economist() +
  labs(
    title = "Total Attrition Distribution"
      )
```


# ------------- Target Analysis ---------------

# Summarize the Attrition Rate  
```{r}
# Gives me a input level breakdown of the target variable
data$Attrition

# Creates a count table (super useful!)
table(data$Attrition)

# Gives me the proportion of each level in the target variable
prop.table(table(data$Attrition)) * 100
```

Frito Lays' Attrition rate is 16.09%. This means that 16.09% of customers have churned.


```{r}
str(data)
```

```{r}
# Unique values in job_level
unique(data$JobLevel)
```

```{r}
null_summary <- sapply(data, function(x) sum(is.na(x)))
null_summary
```

# Univariate Analysis
# Explore each feature individually to understand its distribution and behavior

```{r}
glimpse(data)
```
# Age Distribution
```{r}
ggplot(data, aes(x = Age)) + 
  geom_histogram() +
  geom_density(color = "darkred", size = 1.2) +
  theme_economist() +
  labs(title = "Age Distribution", x = "Age", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```
There is a slight right skew to the Age feature with most of the employees being around the age of 30-40.

# Business Travel Distribution
```{r}
ggplot(data, aes(x = BusinessTravel)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Business Travel Distribution", x = "Business Travel", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )

```

The majority of employees do not travel for business purposes.

# DailyRate Distribution

```{r}
ggplot(data, aes(x = DailyRate)) + 
  geom_histogram(binwidth = 10) +
  theme_economist() +
  labs(title = "Daily Rate Distribution", x = "Daily Rate", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )))

summary(data$DailyRate)
```

The Daily Rate feature appears to be uniformly distributed between 100 and 1500. No particulary useful distribution insights can be derived.

# Department Distribution
```{r}
ggplot(data, aes(x = Department)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Department Distribution", x = "Department", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

The majority of employees work in the Research & Development department, followed by Sales and then Human Resources.

# DistanceFromHome Distribution
```{r}
ggplot(data, aes(x = DistanceFromHome)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Distance From Home Distribution", x = "Distance From Home (miles)", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

There is a heavy concentration of employees living within 10 miles of work, with a gradual decrease in number of employees as distance increases.

# Education Distribution
```{r}
ggplot(data, aes(x = Education)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Education Distribution", x = "Education Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```
 
Education has a normal distribution with the highest concentration of employees centering around level 3.

# EducationField Distribution

```{r}
ggplot(data, aes(x = EducationField)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +
  theme_economist() +
  labs(title = "Education Field Distribution", x = "Education Field", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x  = element_text(margin = margin(t = 12)),
    axis.title.y  = element_text(margin = margin(r = 12)),
    axis.text.x   = element_text(angle = 45, hjust = 1, vjust = 1)   # 👈 rotates labels
  )

```

The majority of employees have a background in Life Sciences, followed by Medical and then Marketing.

# EnvironmentalSatisfaction Distribution

```{r}
ggplot(data, aes(x = EnvironmentSatisfaction)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Environmental Satisfaction Distribution", x = "Environmental Satisfaction Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

The environmental satisfaction seems to show that more employees are satisfied than not.

# Gender Distribution

```{r}
ggplot(data, aes(x = Gender)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Gender Distribution", x = "Gender", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

The gender distribution is showing that there are more Males than females at the company.

# HourlyRate Distribution

```{r}
ggplot(data, aes(x = HourlyRate)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Hourly Rate Distribution", x = "Hourly Rate", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

No useful distribution insights can be derived from the Hourly Rate feature.

# JobInvolvement Distribution
```{r}
ggplot(data, aes(x = JobInvolvement)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Job Involvement Distribution", x = "Job Involvement Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

The job involvement appears to have a the majority of level 3 involvement, with level 4 being the least common.

# JobLevel Distribution
```{r}
ggplot(data, aes(x = JobLevel)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Job Level Distribution", x = "Job Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

The job level distribution shows that the majority of employees are at level 1, with a decreasing number of employees as job level increases.

# MaritalStatus Distribution
```{r}
ggplot(data, aes(x = MaritalStatus)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Marital Status Distribution", x = "Marital Status", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Most employees are single.

# MonthlyIncome Distribution
```{r}
ggplot(data, aes(x = MonthlyIncome)) + 
  geom_histogram(binwidth = 500) +
  theme_economist() +
  labs(title = "Monthly Income Distribution", x = "Monthly Income", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Monthly Income feature appears to be right skewed with most employees earning between $2000 and $8000 per month.

# NumCompaniesWorked Distribution
```{r}
ggplot(data, aes(x = NumCompaniesWorked)) + 
  geom_histogram(binwidth = .8) +
  theme_economist() +
  labs(title = "Number of Companies Worked Distribution", x = "Number of Companies Worked", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The number of companies worked feature appears to be right skewed with most employees having worked at 1-3 companies prior to their current employment.

# OverTime Distribution
```{r}
ggplot(data, aes(x = OverTime)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "OverTime Distribution", x = "OverTime", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Majority of employees do not work overtime.

# PercentSalaryHike Distribution
```{r}
ggplot(data, aes(x = PercentSalaryHike)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Percent Salary Hike Distribution", x = "Percent Salary Hike", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

Percent Salary hike is demonstrating a heavy right skew with most employees receiving between 10-15% salary hikes.

# PerformanceRating Distribution
```{r}
ggplot(data, aes(x = PerformanceRating)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Performance Rating Distribution", x = "Performance Rating", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )

```

Majority of employees have a performance rating of 3, with very few employees receiving a rating of 4.

# RelationshipSatisfaction Distribution
```{r}
ggplot(data, aes(x = RelationshipSatisfaction)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Relationship Satisfaction Distribution", x = "Relationship Satisfaction Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Majority of employees have a relationship satisfaction level of 3 or 4, with level 2 being the least common by a slim margin.

# StandardHours Distribution
```{r}
ggplot(data, aes(x = StandardHours)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Standard Hours Distribution", x = "Standard Hours", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

All employees have standard hours of 80.

# StockOptionLevel Distribution
```{r}
ggplot(data, aes(x = StockOptionLevel)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Stock Option Level Distribution", x = "Stock Option Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Majority of employees have a stock option level of 0, with very few employees having a stock option level of 3.

# TotalWorkingYears Distribution
```{r}
ggplot(data, aes(x = TotalWorkingYears)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Total Working Years Distribution", x = "Total Working Years", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Total Working Years feature appears to be right skewed with most employees having between 0-10 years of total working experience.

# TrainingTimesLastYear Distribution
```{r}
ggplot(data, aes(x = TrainingTimesLastYear)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Training Times Last Year Distribution", x = "Training Times Last Year", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Majority of employees have undergone training 2-3 times in the last year.

# WorkLifeBalance Distribution
```{r}
ggplot(data, aes(x = WorkLifeBalance)) +
  geom_bar() +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.6, size = 4, color = "black"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.02, 0.15))) +  
  theme_economist() +
  labs(title = "Work Life Balance Distribution", x = "Work Life Balance Level", y = "Count") +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
```

Majority of employees have a work life balance level of 3, with level 1 being the least common.

# YearsAtCompany Distribution
```{r}
ggplot(data, aes(x = YearsAtCompany)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Years At Company Distribution", x = "Years At Company", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Years At Company feature appears to be right skewed with most employees having between 0-10 years at the company.

# YearsInCurrentRole Distribution
```{r}
ggplot(data, aes(x = YearsInCurrentRole)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Years In Current Role Distribution", x = "Years In Current Role", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Years In Current Role feature appears to be right skewed with most employees having between 0-7 years in their current role.

# YearsSinceLastPromotion Distribution
```{r}
ggplot(data, aes(x = YearsSinceLastPromotion)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Years Since Last Promotion Distribution", x = "Years Since Last Promotion", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Years Since Last Promotion feature appears to be right skewed with most employees having between 0-3 years since their last promotion.

# YearsWithCurrManager Distribution
```{r}
ggplot(data, aes(x = YearsWithCurrManager)) + 
  geom_histogram(binwidth = 1) +
  theme_economist() +
  labs(title = "Years With Current Manager Distribution", x = "Years With Current Manager", y = "Count") + 
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 12)),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
    )
  )
```

The Years With Current Manager feature appears to be right skewed with most employees having between 0-7 years with their current manager.

# ------------ Bivariate Analysis -------------

