---
title: "Frito Lay: Customer Attrition - Multivariate Analysis"
output: 
  github_document:
    html_preview: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,       
  warning = FALSE,   
  message = FALSE    
)
```

## ------------ Libraries -------------
```{r}
library(tidyverse)
library(caret)
library(e1071)
library(class)
library(scales)
library(ggthemes)
theme_set(theme_economist())
title_format <- function(x) labs(title = x, x = NULL, y = NULL)
```

## ------------ Data Loading -------------
```{r}
# getwd()
data <- read.csv("../data/CaseStudy1-data.csv")
```

## ------------ Multivariate Analysis -------------

Multivariate analysis explores how multiple factors interact to influence employee attrition. By examining combinations of demographic, compensation, and work-related variables together, it reveals deeper relationships—such as how overtime, tenure, and income jointly affect turnover—providing a more complete view of the underlying drivers of employee loss.


--------------------------------------------------------------------------------------------------------------------------------------------------


### Attrition by Department and OverTime

Question:

How does the Attrition rate vary across different Departments and OverTime status? The following heatmap visualizes this relationship:

```{r}
# Calculating attrition rate by Department and OverTime group
df1 <- data %>%
  group_by(Department, OverTime) %>%
  summarise(attr_rate = mean(Attrition == "Yes"), .groups = "drop")

# Creating a heatmap to visualize attrition patterns by Department and OverTime
ggplot(df1, aes(Department, OverTime, fill = attr_rate)) +
  geom_tile(color = "white") +  
  scale_fill_gradient(
    low = "#d8e6f3", 
    high = "#b2182b", 
    labels = percent  
  ) +
  labs(
    title = "Attrition Rate by Department and OverTime",
    fill = "Attrition Rate Percentage"
  ) +
  theme(
    # Center and bold the plot title with slight spacing below
    plot.title = element_text(
      hjust = 0.5,
      face = "bold", 
      margin = margin(b = 10)
    ),
    # Formating legend text and size
    legend.title = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 8),
    legend.key.width = unit(1, "cm"),
    legend.key.height = unit(0.4, "cm"),
    # Adding padding around axis titles for readability
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12)),
  )

```

This heatmap illustrates how attrition rates vary across departments and overtime status. The darkest red shading represents the highest attrition percentages. Employees in the Sales department who work overtime experience the highest attrition rate, exceeding 40%. In contrast, Research & Development employees with no overtime show the lowest attrition, suggesting that work-life balance plays a key role in retention. Across all departments, overtime consistently correlates with higher attrition, but the impact is most pronounced in Sales—indicating that targeted workload management and incentive strategies in Sales could potentially yield meaningful reductions in turnover.


--------------------------------------------------------------------------------------------------------------------------------------------------


### Attrition by Job Role and Education Field

Question:

How does the attrition rate vary across different Job Roles and Education Fields?

```{r}
# Calculating attrition rate by Job Role and Education Field
df2 <- data %>%
  group_by(JobRole, EducationField) %>%
  summarise(attr_rate = mean(Attrition == "Yes"), .groups = "drop")

# Plotting a Heatmap to visualize 
ggplot(df2, aes(EducationField, JobRole, fill = attr_rate)) +
  geom_tile(color = "white") +   
  scale_fill_gradient(
    low  = "#d8e6f3",            
    high = "#b2182b",            
    labels = percent             
  ) +
  labs(
    title = "Attrition Rate by Job Role and Education Field",
    fill  = "Attrition Rate Percentage"
  ) +
  theme(
    # Tweaking plot appearance
    plot.title = element_text(hjust = 0.5, face = "bold", margin = margin(b = 10)),
    legend.position = "top",
    legend.justification = "center",
    legend.title = element_text(size = 9, face = "bold"),
    legend.text  = element_text(size = 8),
    legend.key.width  = unit(1, "cm"),
    legend.key.height = unit(0.4, "cm"),
    # Improve x-axis label readability
    axis.text.x  = element_text(angle = 45, hjust = 1, vjust = 1, size = 9),
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12)),
    # Balancing white space left/right (long y labels widen the left side)
    plot.margin  = margin(t = 5, r = 75, b = 5, l = 5)
  )

```

This visualization highlights that attrition risk is not evenly distributed across job roles or educational backgrounds. Employees in sales-related roles—particularly Sales Representatives with Medical or Technical degrees—show the highest turnover rates. Conversely, employees in research and management positions display significantly lower attrition, indicating stronger role alignment and job stability. These insights suggest that retention efforts may need to focus on improving engagement, career development, and growth opportunities for technically trained employees in high-turnover sales positions.


--------------------------------------------------------------------------------------------------------------------------------------------------


### Income Distribution for Sales Employees by Attrition

Question:

Do income levels differ significantly between employees who stay and those who leave within the Sales department?

```{r}
# Boxplot to compare Monthly Income distribution for Sales employees by Attrition status
ggplot(data %>% filter(Department == "Sales"),
       aes(x = Attrition, y = MonthlyIncome, fill = Attrition)) +
  # Creating notched boxplots to visualize distribution and median confidence intervals
  geom_boxplot(
    color = "black",
    outlier.shape = 16,
    outlier.color = "black",
    outlier.size = 2,
    notch = TRUE,
    notchwidth = 0.9
  ) +
  scale_fill_manual(
    values = c("Yes" = "#b2182b", "No" = "#2166ac")
  ) +
  # Format y-axis values as dollar amounts
  scale_y_continuous(labels = dollar_format(prefix = "$", big.mark = ",")) +

  # Add median labels to each box for clarity
  stat_summary(
    fun = median,
    geom = "label",
    aes(label = paste0("Median: $", format(round(..y.., 0), big.mark = ","))),
    color = "black",
    fill = "white",
    fontface = "bold",
    size = 4,
    label.size = 0.1,
    vjust = -0.5
  ) +
  labs(
    title = "Income Distribution for Sales Employees by Attrition",
    x = "Attrition",
    y = "Monthly Income"
  ) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      margin = margin(b = 10)
    ),
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 12)),
    axis.title.y = element_text(margin = margin(r = 12))
  )

```

This boxplot provides a clear comparison of monthly income between Sales employees who stayed and those who left the company. The median income for employees who did not leave is approximately $5,900, whereas those who left have a notably lower median of $4,700. These findings reinforce an idea that lower compensation may be a significant driver of attrition in the Sales department, highlighting an opportunity for Frito Lay to improve retention through targeted pay adjustments or incentive structures.


--------------------------------------------------------------------------------------------------------------------------------------------------


